# Reproducible Research: Peer Assessment 1
========================================================

```{r, echo=FALSE}
## Setting the working directory
setwd("D:\\#FORMACION\\Coursera\\05 - Reproducible Research\\PeerA1\\RepData_PeerAssessment1")
date <- date()
```

## Loading and preprocessing the data
Step 1: Read the data: 

```{r}
#setInternet2(TRUE)
#fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#download.file(fileUrl, destfile="repdata_Fdata_Factivity.zip", method="auto")
Factivity <- read.csv(unz("repdata_Fdata_Factivity.zip","activity.csv"))
unlink("activity.csv")
```

Date file download: 
```{r} 
date
```

## What is mean total number of steps taken per day?

```{r} 
# Histogram of the total number of steps taken each day
aggSum <- aggregate (Factivity$steps, by=list(as.factor(Factivity$date)),FUN=sum,rm.na=TRUE)
names(aggSum) <- c("date","steps")
hist(aggSum$steps,main="Steps Histogram",col="blue",xlab="Steps")
# Mean and Median total number of steps taken per day
mean(aggSum$steps, na.rm=TRUE)
median(aggSum$steps, na.rm=TRUE)
```

## What is the average daily activity pattern?

```{r} 
## Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

avgstepinterv <- aggregate (Factivity$steps, by=list(as.factor(Factivity$interval)),FUN=mean, na.rm=TRUE)
names(avgstepinterv) <- c("interval","mean.steps")

library(ggplot2)
ggplot(avgstepinterv, aes(x=as.factor(interval), y=mean.steps))+
geom_line(aes(group=1), colour="#000099") +  # Blue lines
geom_point(size=1, colour="#CC0000") +        # Red dots
labs(x="Hour Interval")+labs(y="Average number of Steps") +
        labs(title="Daily Activity Pattern")
#5-minute interval that contains the maximum number of steps on average across all the days in the dataset,     
avgstepinterv[avgstepinterv$mean.steps==max(avgstepinterv$mean.steps),]
```



## Imputing missing values

```{r} 
# Total number of missing values in the dataset 
nrow(Factivity[is.na(Factivity$steps),])

#Strategy for filling in all of the missing values in the dataset: 
#Use the mean for that 5-minute interval, etc.

#New dataset that is equal to the original dataset but with the missing data filled in.
total <- merge(Factivity,avgstepinterv,by="interval")
total <- total[order(total$date,total$interval),]
total$steps1 <- ifelse(is.na(total$steps),total$mean.steps,total$steps)

#Histogram of the total number of steps taken each day (replacing NA values)
aggSum1 <- aggregate (total$steps1, by=list(as.factor(Factivity$date)),FUN=sum,rm.na=TRUE)
names(aggSum1) <- c("date","steps")
hist(aggSum1$steps,main="Steps Histogram",col="blue",xlab="Steps")
# Mean and Median total number of steps taken per day
mean(aggSum1$steps, na.rm=TRUE)
median(aggSum1$steps, na.rm=TRUE)


```

These values nearly does not differ from the estimates from the first part of the assignment. Imputing missing data on the estimates of the total daily number of steps, results in higher median value.


## Are there differences in activity patterns between weekdays and weekends?

## What is the average daily activity pattern?

```{r} 
avgsteps.interval <- aggregate (Factivity$steps, by=list(as.factor(Factivity$interval)),FUN=mean, na.rm=TRUE)
names(avgsteps.interval) <- c("interval","mean.steps")
hinterv <- as.factor(avgsteps.interval$interval)
plot(hinterv, avgsteps.interval$mean.steps, type="s", xlab="Hour Interval", ylab="Average number of Steps",main="Average Daily Activity Pattern")

 avgsteps.interval[avgsteps.interval$mean.steps==max(avgsteps.interval$mean.steps),]
```
